\textsf{Are large language models superhuman chemists?}

### 2024-03-21 src/tex/ms.tex
make folder for sections
update text for question selection

@@

@@

\begin{figure}[htb]
    \centering
    \includegraphics{figures/human_subset_performance.pdf}
    \caption{\textbf{Performance of models and humans on \chembench.} The figure shows the percentage of questions that the models answered completely correctly. We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance. 
    Since the humans did not answer all the questions, this plot is based on the subset of questions that most humans answered.
    The metric we use here is harsh as it only considers a question answered completely correctly or completely incorrectly, partially correct answers are also considered as incorrect.
    \Cref{fig:barplot_all_correct_all_questions} gives an overview of the performance of various models on the entire corpus.
    }
    \label{fig:human_vs_models_bar}
    \script{plot_overview_performance_plot.py}
\end{figure}

@@

\begin{figure}[htb]
    \centering
    \includegraphics{figures/all_questions_models_completely_correct_radar_human.pdf}
    \caption{\textbf{Performance of the models and humans on the different topics of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus. The performance is measured as the fraction of questions that were answered completely correctly by the models.
    The best score for every dimension is one (all questions answered correctly) and the worst score is zero (no question answered completely correctly). A larger colored area indicates a better performance.
    This figure shows the performance on the subset of questions that was answered by humans, too. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
    }
    \label{fig:all_questions_models_completely_correct_radar_human}
    \script{analyze_model_reports.py}
\end{figure}

@@

% GFK performance -> would the models obtain a chemical license 
% Interesting twist perhaps: the performance on the university textbooks or exams is better than on the ones we automatically constructed
% in contrast to LLM the human performance is also less spiky

@@

@@

\paragraph{Question selection} \label{sec:subset-selection}
Since we anticipated that we will not be able to collect enough responses for every question to allow for a meaningful statistical analysis, we decided on showing a relevant subset of all questions to the human scorers.
For selecting the subset, we decided on addressing two questions:
\begin{itemize}
    \item Are the questions for which the models scored poorly just too difficult or unanswerable?
    \item Are there areas in which the performance of humans is very different from the ones of the models?
\end{itemize}
To answer the first question we selected X questions that all \glspl{llm} (model names) from an initial scoring round did not answer correctly.
From those we picked X diverse one using greedy MaxMin sampling on the embeddings on the questions computed using BART (see below).

>>>>

% make the app somewhat more prominent

@@

% emphasize that the topics are more balanced in this subset

@@

\begin{figure}[htb]
    \centering
    \includegraphics{figures/human_subset_performance.pdf}
    \caption{\textbf{Performance of models and humans on the \enquote{tiny} subset of the  \chembench corpus.} The figure shows the percentage of questions that the models answered completely correctly. We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance. 
    Since the humans did not answer all the questions, this plot is based on the subset of questions that most humans answered.
    The metric we use here is harsh as it only considers a question answered completely correctly or completely incorrectly, partially correct answers are also considered as incorrect.
    \Cref{fig:barplot_all_correct_all_questions} gives an overview of the performance of various models on the entire corpus.
    }
    \label{fig:human_vs_models_bar}
    \script{plot_overview_performance_plot.py}
\end{figure}

@@

\begin{figure}[htb]
    \centering
    \includegraphics{figures/all_questions_models_completely_correct_radar_human.pdf}
    \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus. The performance is measured as the fraction of questions that were answered completely correctly by the models.
    The best score for every dimension is one (all questions answered correctly) and the worst score is zero (no question answered completely correctly). A larger colored area indicates a better performance.
    This figure shows the performance on the subset of questions that was answered by humans, too. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
    }
    \label{fig:all_questions_models_completely_correct_radar_human}
    \script{analyze_model_reports.py}
\end{figure}

@@

% GFK performance -> would the models obtain a chemical license 
% Interesting twist perhaps: the performance on the university textbooks or exams is better than on the ones we automatically constructed
% in contrast to LLM the human performance is also less spiky
% emphasise that the exact ranking depends on the subset of questions we look at

@@

\input{sections/parse_check_desc.tex}

@@

\paragraph{Question selection} \label{sec:subset-selection}
Since we anticipated that we will not be able to collect enough responses for every question to allow for a meaningful statistical analysis, we decided on showing a relevant subset of all questions to the human scorers.
For selecting the subset, we decided on addressing two questions:
\begin{itemize}
    \item Are the questions for which the models scored poorly just too difficult or unanswerable?
    \item Are there areas in which the performance of humans is very different from the ones of the models?
\end{itemize}
To answer the first question we selected up to 13 questions per source that all \glspl{llm} (model names) from an initial scoring round did not answer correctly.
In addition, we picked 100 diverse one using greedy MaxMin sampling on the embeddings on the questions computed using BART and added five random questions about the number of \gls{nmr} peaks and the point group of molecules.
